### 2008-07-24 doc/storage_system/arc1-storage-documentation.tex
almost done with the technical documentation of the ARC1 storage

\documentclass{book}
\usepackage{graphicx}                              %for PNG images (pdflatex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {Zsombor Nagy},
  pdftitle = {Documentation of the ARC1 storage system},
  pdfsubject = {Paper subject},
  pdfkeywords = {Paper,keyword,comma-separated},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~2~}
\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
%----- DON'T CHANGE HEADER MATTER
\begin{document}
\def\today{\number\day/\number\month/\number\year}

% \begin{titlepage}
% 
% \begin{tabular}{rl}
% \resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
% &\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
% \end{tabular}
% 
% \hrulefill
% 
% %-------- Change this to NORDUGRID-XXXXXXX-NN
% 
% {\raggedleft NORDUGRID-XXXXXXX-NN\par}
% 
% {\raggedleft \today\par}
% 
% \vspace*{2cm}
% 
% %%%%---- The title ----
% {\centering \textsc{\Large Documentation of the ARC1 storage system}\Large \par}
% \vspace*{0.5cm}
%     
% %%%%---- A subtitle, if necessary ----
% {\centering \textit{\large First prototype status and plans}\large \par}
%     
% \vspace*{1.5cm}
% %%%%---- A list of authors ----
%     {\centering \large Zsombor Nagy\footnote{zsombor@niif.hu} \large \par}
%     
% \end{titlepage}
% 
% \tableofcontents                          %Comment if use article style
% \newpage
\chapter{Design Overview}
\label{sec:overview}

% NorduGrid~\cite{nordugrid} Advanced Resource Connector (ARC)
% \textit{middleware}\index{middleware} uses \GT~2 by Globus.
% 
% In expressions, the following operands are allowed:
% \begin{shaded}
%   \verb#=   !=   >   <   >=   <=#
% \end{shaded}
% 
% \begin{framed}
%   Examples of URLs are:\\
%   \\
%   \verb#http://grid.domain.org/dir/script.sh#\\
%   \verb#gsiftp://grid.domain.org:2811;threads=10/dir/input_12378.dat#\\
%   \verb#ldap://grid.domain.org:389/lc=collection1,rc=Nordugrid,dc=nordugrid,dc=org#\\
%   \verb#rc://grid.domain.org/lc=collection1,rc=Nordugrid,dc=nordugrid,dc=org/zebra/f1.zebra#
%   \verb#file:///home/auser/griddir/steer.cra#\\
% \end{framed}

The ARC1 storage system is a distributed system for storing replicated \emph{files} on several file storage nodes and manage them in a global namespace. The files can be grouped into collections, and a \emph{collection} can contain sub-collections and sub-sub-collections in any depth. There is a dedicated \emph{root collection} to gather all collections to the global namespace. This hierarchy of collections and files can be referenced using \emph{Logical Names} (LN). The users can use this global namespace  as they were using a local filesystem. Files can be transfered by multiple transfer protocols, and the client side tools hide this from the user. The replicas of the files are stored on different storage nodes. A storage node should have one or more data transfer service running (e.g. HTTP(S), FTP(S), GridFTP, ByteIO, etc.), and one of the ARC1 storage system’s services is needed to manage each storage node and integrating it into the system. And there are plans to provide a way to integrate third-party storage solutions into the namespace of the ARC1 storage system. The main services are shown on Figure~\ref{fig:services}.

\begin{figure}[ht]
\centering{{\scalebox{0.9}{\includegraphics{arc1-storage-services.pdf}}}
\caption{\label{fig:services}The components of the ARC1 storage;
A-H: the \textbf{A-Hash}, which is a currently centralized, later distributed database;
L: the \textbf{Librarian}, which stores the metadata and hierarchy of collections and files, the location of replicas, and health data of the Shepherd services, using the A-Hash as database;
B: the \textbf{Bartender}, which provides a high-level interface for the users and for other services;
S: the \textbf{Shepherd}, which provides a simple interface for storing files on storage nodes.
} }
\end{figure}

@@

There are a number of IDs used in the ARC storage system, such as:
\begin{itemize}
    \item Each file and collection has a globally unique ID called \textbf{GUID}.
    \item The files and collections are organized into a global hierarchical namespace and can be referred to using paths of this namespace called \textbf{Logical Names (LN)}.
    \item Each service has a unique \textbf{serviceID} which can be used to get an endpoint reference from the information system. We need an endpoint reference which is an address (URL) which we could connect to.
    \item The Shepherds in the system identify their files using a \textbf{referenceID}.
    \item The \textbf{location} of a replica consists of two IDs: the ID of the Shepherd and the ID of the file within the Shepherd: (serviceID, referenceID).
\end{itemize}

@@

\section{GUIDs and Logical Names (LN)} % (fold)
\label{sec:GUIDsLNs}

The syntax of Logical Names (LN): \verb#/[path]# or \verb#<GUID>[/<path>]# where [...] indicates optional parts.

@@

\begin{figure}[ht]
\centering{{\scalebox{0.7}{\includegraphics{arc1-storage-namespace.pdf}}}
\caption{\label{fig:namespace}Example of the global namespace's hierarchy} }
\end{figure}

Example on Figure~\ref{fig:namespace}: if we have a collection with GUID \verb#1234#, and there is a collection called \verb#green# in it, and in \verb#green# there is another collection called \verb#orange#, and in \verb#orange# there is a file called \verb#huge#, then we can refer to this file as \verb#/green/orange/huge#, if we know the GUID of the starting collection, so let's prefix the path with it: \verb#1234/green/orange/huge#. This is the Logical Name of that file. If there is a well-known system-wide root collection (its GUID could be e.g. \verb#0#), then if a LN starts with no GUID prefix, it is implicitly prefixed with the GUID of this well-known root collection, e.g. \verb#/why/blue# means \verb#0/why/blue#.

If a client wants to find the file called \verb#/why/blue#, the system knows where to start the search: the GUID of the root collection. The root collection knows the GUID of \verb#why#, and the (sub-)collection \verb#why# knows the GUID of \verb#blue#. If the GUID of this file is \verb#5678#, and somebody makes another entry in collection \verb#/why# (= \verb#0/why#) with name \verb#red# and GUID \verb#5678#, then the \verb#/why/red# LN points to the same file as \verb#/why/blue#, so it's a hard link. To count the references (the number of hardlinks) of a file or collection, the GUID of the parent collection(s) is stored as a metadata for each file and collection.

Each VO should create a VO-wide root collection, and put it in the generic root collection, e.g. if a VO called \verb#vo1# creates a collection called \verb#vo1# as a sub-collection of the root collection (which has the GUID \verb#0#), then it can be referred as \verb#0/vo1# or just \verb#/vo1#. Then this VO can create some files, and put them in this \verb#/vo1# collection, e.g. \verb#/vo1/file1#, etc. Or sub-collections, e.g. \verb#/vo1/col1#, \verb#/vo1/col2/file3#, etc. For this the VO does not need the install any service. These files and collections can be created using a Bartender service.

@@

\section{The Bartenders} % (fold)
\label{sec:the_bartenders}
Clients can access the storage system through a Bartender service. If a client wants to create a collection, upload or download a file, the first step is to connect a Bartender. The Bartender then asks a Librarian to resolve Logical Names and return metadata, then initiates file transfers on some storage node by asking the Shepherd service on the node, then returns the transfer URL to the client, which allows the client to directly transfer from/to the storage node. So the data transfer itself is not going through the Bartender, it is performed over a direct link between a storage node and the client.

@@

\section{The Librarians} % (fold)
\label{sec:the_librarians}
The Librarian is capable of managing the hierarchy and metadata of files and collections, and health information of the Shepherd services. Each file and collection in the Librarian has a globally unique ID (GUID). A collection contains files and other collections, and each of these entries has a name unique within the collection very much like entries in a usual directory on a local filesystem. Besides files and collections the Librarian stores a third type of entries called Mount Points which are references to external services creating the capability to mount the namespace of third-party storages to our global namespace and make the files on a third-party storage available through the interface of the ARC storage system.

@@

\section{The A-Hash} % (fold)
\label{sec:the_a_hash}
The A-Hash is a currently centralized, but later distributed service capable of consistently storing objects containing property-value pairs organized in sections. All metadata about files and collections are stored in the A-Hash, and some information about Shepherd services is stored in it as well.

@@

\section{The Shepherds} % (fold)
\label{sec:the_shepherds}
When a new file is put into the system the number of needed replicas is given for the file. The file replicas are stored on different storage nodes, for each storage node there is Shepherd service which provides the interface for initiating file transfer.

The file-naming used by a Shepherd has nothing to do with the the hierarchy of collections, or Logical Names. When a replica is given to a Shepherd it gets an ID which refers to it within that Shepherd. Each Shepherd has a unique ID itself, so with these IDs the replica can be unambiguously referenced, this is called a Location. The namespace of these Locations has nothing to do with the namespace of GUIDs or the namespace of Logical Names. It consists of two IDs: the ID of the Shepherd and the ID of the file within the Shepherd: \verb!(serviceID, referenceID)!

@@

If a Shepherd finds out that a file is missing or has a bad checksum, it reports that the file is \verb#invalid# to the Librarian immediately, and the Librarian alters the state of the given replica of the file.

The Shepherd tries to recover its replica by downloading it from another Shepherd. In order to do this the Shepherd contacts a Bartender and asks for the file. The Bartender chooses a valid replica, initiate file transfer by a Shepherd having a valid replica, and returns the TURL to the Shepherd with the invalid replica. The Shepherd with the invalid replica downloads the file from the other Shepherd, and if everything is OK, signals to the Librarian that the replica is \verb!alive! again.

@@

We want to download a file about which we know that it is somewhere in our home collection on the storage (see Figure~\ref{fig:downloading}). The LN of our home collection is e.g. \verb#/ourvo/users/we#. We can get a list of entries in this collection from any Bartender.

\begin{enumerate}
\item We need to find a Bartender. Maybe we have a cached list of recently used Bartenders or we can get one from the information system. When we have an endpoint reference of a Bartender, we could call its list method with the LN \verb#/ourvo/users/we#.
\item The Bartender has to find a Librarian service, again using its cache of recently used Librarian services or get a new one from the information system. When rhe Bartender has an endpoint reference of a Librarian service, it could ask the Librarian to traverse the LN \verb#/ourvo/users/we#.
\item The Librarian needs the A-Hash service to access the stored data, when it has the endpoint reference of the A-Hash service, it could get the information about the root collection, which contains the GUID of the \verb#ourvo# sub-collection. Then the Librarian gets the entries of this \verb#ourvo# collection, and in it it can find the GUID of \verb#users#, and in the entries of \verb#users# there is the GUID of \verb#we#, which the Librarian returns to the Bartender with all the metadata.
\item The Bartender now has the GUID and the metadata of the collection \verb#/ourvo/users/we#, including the list of its entries. This is returned to us.
\item So we get the list of our \verb#/ourvo/users/we# collection, and now we realize that the file we want has the LN \verb#/ourvo/users/we/thefilewewant# and we know the GUID of it as well: e.g. \verb#a4b2e#. (Of course we know the GUID of the \verb#/ourvo/users/we# collection too, which is e.g. \verb#13245# and using this we could refer to our file as \verb#13245/thefilewewant# which means the entry called \verb#thefilewewant# in the collection with a GUID \verb#13245#.) We connect a Bartender again (the same one or maybe another one) to get the file with any of these LNs, the \verb#a4b2e# is the fastest solution because the Bartender need not to look up the whole LN again in the Librarian, a well-written client API should use this. With the get request we give the Bartender the list of transfer protocols we are able to use.
\item The Bartender contacts the Librarian to get the locations of the replicas of this file.
\item The Librarian returns all the metadata of the requested LN.
\item The Bartender chooses a ‘location’ which consists of the ID of a Shepherd, and the referenceID of the file within the Shepherd. Using the information system or its local cache it could get the endpoint reference of the Shepherd. The Bartender initiates a transfer by the Shepherd, if the Shepherd supports one of the transfer protocols we give, it can create a transfer URL (TURL) with a protocol we can download.
\item The Shepherd returns the TURL to the Bartender.
\item The Bartender returns the TURL to us.
\item Now we have a TURL from which we can download it.
\end{enumerate}

@@

We have a file on our local disk we want to upload to a collection called \verb#/ourvo/common/docs#. (See Figure~\ref{fig:uploading}.)
\begin{enumerate}
\item We contact a Bartender to put the file, we give the size and checksum and other metadata, and the transfer protocols we want to use. And of course we give the Logical Name we want to be the name of the file, which in this case will be \verb#/ourvo/common/docs/proposal.pdf#
\item The Bartender ask a Librarian to traverse this LN.
\item If the Librarian can traverse the whole LN then this LN is already exists, but if the LN is still available and the parent exists, the Librarian’s response contains the GUID and metadata of the parent collection.
\item Then the Bartender creates a new file entry within the Librarian with all the information we gave. 
\item The Librarian returns the GUID of this new entry.
\item Then the Bartender add the name \verb#proposal.pdf# and the new GUID to the collection \verb#/ourvo/common/docs# and from now on there will be a valid LN \verb#/ourvo/common/docs/proposal.pdf# which points to a file which has no replica at all. If someone tried to download the file called \verb#/ourvo/common/docs/proposal.pdf# now, would get an error message ‘try again later’.
\item The Bartender (using the information system) chooses a Shepherd and gets its endpoint reference. Then the Bartender initiates uploading of the file to the Shepherd: the request includes the size and checksum of the file, the GUID, and the protocols we are able to use.
\item The Shepherd creates a transfer URL and a referenceID for this file and registers the GUID of the file in its own database and reports to the Librarian that there is a new replica with state \verb#creating#. The Librarian gets the message from the Shepherd and creates a new entry in the locations list of the given file with the serviceID and the referenceID the Shepherd have just reported. If someone tries to download this file now, still gets a ‘try again later’ error message, because this new replica is still not \verb#alive#.
\item The Shepherd returns the the TURL to the Bartender.
\item The Bartender returns the TURL to us.
\item Then we can upload the file to this TURL.
\item The Shepherd detects that the file is arrived and reports the change of state to \verb#alive# to the  Librarian who alters the state in the given file-entry. At this point the file has only one replica. 
\end{enumerate}
\begin{itemize}
\item The Shepherd periodically checks the Librarian if this is less than the needed replica number, and if it is then it initiates creating a new replica by a Bartender.
\item The Bartender chooses another Shepherd, initiates the transfer then returns the TURL to the  first Shepherd which uploads the file to the new Shepherd.
\item All the Shepherds check periodically whether their files have enough replica, and if any of them find that there is more replica needed, it initiates creating a new. If more than one Shepperd of course could cause that there will be more replicas than needed. If a Shepherd finds out that a file has more replicas than needed it notifies a Bartender about it.
\item The Bartender ask the Librarian about all Shepherds this file has replicas on, flags this file as ‘removing a replica’ which prevents other Bartenders to remove an other replica accidentally, then make a decision of which one is to be removed, then contacts the chosen Shepherd and asks it to remove the replica. The Shepherd then notify the Librarian, and the Librarian removes the replica, and removes the flag ‘removing a replica’ as well.
\item If we cannot upload the file to the given TURL for some reason, we should remove the file entry from the collection, or we should call \verb#addReplica# to get a new TURL without removing and recreating the file.
\end{itemize}

@@

\begin{figure}[ht]
\centering{{\scalebox{0.7}{\includegraphics{arc1-storage-removing.pdf}}}
\caption{\label{fig:removing}Removing a file} }
\end{figure}
\begin{enumerate}
    \item If we want to remove a file, we should connect to a Bartender with the LN of the file we want to remove.
    \item The Bartender asks the Librarian to traverse the LN and return the list of parent collections of this file.
    \item The Librarian returns the data.
    \item If the file has only parent, it asks the Librarian to remove the file entry itself, and the entry from the parent collection. If the file has more parent collections (hardlinks) the Bartender only removes the entry from the parent collection and removes the parent from the list of parent  collections of this file.
    \item Next time the Shepherd does its periodic check, it asks the Librarian about each of its stored replicas, and finds out that one of them no longer exists, so it removes the replica from the storage node.
\end{enumerate}

@@

The current version of the prototype has no information system and no security, these are soon to be integrated to the system.
The information system is needed to discover services, and to translate serviceIDs to endpoint references (URLs). Currently the URLs are written in the configuration files, and the Shepherd services are reporting their URLs to the Librarian, so the Bartender could ask for all alive Shepherds.

>>>>

\documentclass{book}
\usepackage{graphicx}                              %for PNG images (pdflatex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {Zsombor Nagy},
  pdftitle = {Documentation of the ARC1 storage system},
  pdfsubject = {Paper subject},
  pdfkeywords = {Paper,keyword,comma-separated},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
%----- DON'T CHANGE HEADER MATTER
\begin{document}
\def\today{\number\day/\number\month/\number\year}

\begin{titlepage}

\begin{tabular}{rl}
\resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
&\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
\end{tabular}

\hrulefill

%-------- Change this to NORDUGRID-XXXXXXX-NN

{\raggedleft NORDUGRID-XXXXXXX-NN\par}

{\raggedleft \today\par}

\vspace*{2cm}

%%%%---- The title ----
{\centering \textsc{\Large Documentation of the ARC1 storage system}\Large \par}
\vspace*{0.5cm}

%%%%---- A subtitle, if necessary ----
{\centering \textit{\large First prototype status and plans}\large \par}

\vspace*{1.5cm}
%%%%---- A list of authors ----
    {\centering \large Zsombor Nagy\footnote{zsombor@niif.hu} \large \par}

\end{titlepage}

\tableofcontents                          %Comment if use article style
\newpage

\renewcommand{\thefootnote}{\arabic{footnote}}

\chapter{Design Overview}
\label{sec:overview}

The ARC1 storage system is a distributed system for storing replicated \emph{file}s on several file storage nodes and manage them in a global namespace. The files can be grouped into \emph{collection}s, and a collection can contain sub-collections and sub-sub-collections in any depth. There is a dedicated \emph{root collection} to gather all collections to the global namespace. This hierarchy of collections and files can be referenced using \emph{Logical Name}s (\emph{LN}s). The users can use this global namespace  as they were using a local filesystem. Files can be transfered by multiple transfer protocols, and the client side tools hide this from the user. The replicas of the files are stored on different storage nodes. A storage node should have one or more data transfer service running (e.g. HTTP(S), FTP(S), GridFTP, ByteIO\footnote{OGSA ByteIO Working Group (BYTEIO-WG), https://forge.gridforum.org/projects/byteio-wg/}, etc.), and one of the services of the ARC1 storage system is needed to manage each storage node and to integrate it into the system. And there are plans to provide a way to integrate third-party storage solutions into the namespace of the ARC1 storage system. The main services are shown on Figure~\ref{fig:services}.

\begin{figure}[ht]
\centering{{\scalebox{0.9}{\includegraphics{arc1-storage-services.pdf}}}
\caption{\label{fig:services}The components of the ARC1 storage;
\emph{A-H}: the \textbf{A-Hash}, which is a currently centralized, later distributed database;
\emph{L}: the \textbf{Librarian}, which stores the metadata and hierarchy of collections and files, the location of replicas, and health data of the Shepherd services, using the A-Hash as database;
\emph{B}: the \textbf{Bartender}, which provides a high-level interface for the users and for other services;
\emph{S}: the \textbf{Shepherd}, which provides a simple interface for storing files on storage nodes.
} }
\end{figure}

@@

There are a number of IDs used in the ARC storage system, such as:
\begin{itemize}
    \item Each file and collection has a globally unique ID called \textbf{GUID}.
    \item The files and collections are organized into a global hierarchical namespace and can be referred to using paths of this namespace called \textbf{Logical Name}s (\textbf{LN}s).
    \item Each service has a unique \textbf{serviceID} which can be used to get an endpoint reference from the information system. We need an endpoint reference which is an address (URL) which we could connect to.
    \item The Shepherds in the system identify their files using a \textbf{referenceID}.
    \item The \textbf{location} of a replica consists of two IDs: the ID of the Shepherd and the ID of the file within the Shepherd: (serviceID, referenceID).
\end{itemize}

@@

\section{GUIDs and Logical Names (LNs)} % (fold)
\label{sec:GUIDsLNs}

The syntax of Logical Names is \verb#/[path]# or \verb#<GUID>[/<path>]# where [...] indicates optional parts.

@@

\begin{figure}[ht]
\centering{{\scalebox{0.7}{\includegraphics{arc1-storage-namespace.pdf}}}
\caption{\label{fig:namespace}Example of the hierarchy of the global namespace} }
\end{figure}

Example on Figure~\ref{fig:namespace}: if we have a collection with GUID \verb#1234#, and there is a collection called \verb#green# in it, and in \verb#green# there is another collection called \verb#orange#, and in \verb#orange# there is a file called \verb#huge#, then we can refer to this file as \verb#/green/orange/huge# if we know the GUID of the starting collection, so let's prefix the path with it: \verb#1234/green/orange/huge#. This is the Logical Name of that file. If there is a well-known system-wide root collection (its GUID could be e.g. \verb#0#), then if a LN starts with no GUID prefix, it is implicitly prefixed with the GUID of this well-known root collection, e.g. \verb#/why/blue# means \verb#0/why/blue#.

If a client wants to find the file called \verb#/why/blue#, the system knows where to start the search: the GUID of the root collection. The root collection knows the GUID of \verb#why#, and the (sub-)collection \verb#why# knows the GUID of \verb#blue#. If the GUID of this file is \verb#5678#, and somebody makes another entry in collection \verb#/why# (= \verb#0/why#) with name \verb#red# and GUID \verb#5678#, then the \verb#/why/red# LN points to the same file as \verb#/why/blue#, so it's a hard link. To count the references (the number of hard links) of a file or collection, the GUID of the parent collection(s) is stored as a metadata for each file and collection.

Each VO\footnote{Virtual Organization, http://www.globus.org/grid\_software/security/voms.php} should create a VO-wide root collection, and put it in the generic root collection, e.g. if a VO called \verb#vo1# creates a collection called \verb#vo1# as a sub-collection of the root collection (which has the GUID \verb#0#), then it can be referred as \verb#0/vo1# or just \verb#/vo1#. Then this VO can create some files, and put them in this \verb#/vo1# collection, e.g. \verb#/vo1/file1#, etc. Or sub-collections, e.g. \verb#/vo1/col1#, \verb#/vo1/col2/file3#, etc. For this the VO does not need the install any service. These files and collections can be created using a Bartender service.

@@

\section{The Bartenders} % (fold)
\label{sec:the_bartenders}
Clients can access the storage system through a Bartender service. If a client wants to create a collection, upload or download a file, the first step is to connect to a Bartender. The Bartender then asks a Librarian to resolve Logical Names and return metadata, then initiates file transfers on some storage node by asking the Shepherd service of that node, then returns the transfer URL (\emph{TURL}) to the client, which allows the client to directly transfer from/to the storage node. So the data transfer itself is not going through the Bartender, it is performed over a direct link between a storage node and the client.

@@

\section{The Librarians} % (fold)
\label{sec:the_librarians}
The Librarian is capable of managing the hierarchy and metadata of files and collections, and health information of the Shepherd services. Each file and collection in the Librarian has a globally unique ID (GUID). A collection contains files and other collections, and each of these entries has a name unique within the collection very much like entries in a usual directory on a local filesystem. Besides files and collections the Librarian stores a third type of entries called Mount Points which are references to external services which will be in the future used to mount the namespace of third-party storages to our global namespace and make the files on a third-party storage available through the interface of the ARC storage system.

@@

\section{The A-Hash} % (fold)
\label{sec:the_a_hash}
The A-Hash is a currently centralized, but later distributed service capable of consistently storing objects containing property-value pairs organized in sections. All metadata about files and collections are stored in the A-Hash, and some information about Shepherd services is stored in it as well. The A-Hash itself does not interpret the data, it just stores tuples of strings.

@@

\section{The Shepherds} % (fold)
\label{sec:the_shepherds}
When a new file is put into the system the number of needed replicas is given for the file. The file replicas are stored on different storage nodes, for each storage node there is Shepherd service which manages the storage node, reports its health state to the Librarian and provides the interface for initiating file transfer.

The file-naming used by a Shepherd has nothing to do with the the hierarchy of collections, or Logical Names. When a new replica upload is initiated, the Shepherd generates an ID which refers to it within that Shepherd. Each Shepherd has a unique ID itself, so with these IDs the replica can be unambiguously referenced, this is called a \emph{Location}. The namespace of these Locations has nothing to do with the namespace of GUIDs or the namespace of Logical Names. It consists of two IDs: the ID of the Shepherd and the ID of the file within the Shepherd: \verb!(serviceID, referenceID)!

@@

If a Shepherd finds out that a file is missing or has a bad checksum, it reports that the file is \verb#invalid# to the Librarian immediately, and the Librarian alters the state of the given replica of the file. The Shepherd tries to recover its replica by downloading it from another Shepherd. In order to do this the Shepherd contacts a Bartender and asks for the file. The Bartender chooses a valid replica, initiate file transfer by a Shepherd having a valid replica, and returns the TURL to the Shepherd with the invalid replica. The Shepherd with the invalid replica downloads the file from the other Shepherd, and if everything is OK, signals to the Librarian that the replica is \verb!alive! again.

@@

We want to download a file about which we know that it is somewhere in our home collection on the storage (see Figure~\ref{fig:downloading}). The LN of our home collection is e.g. \verb#/ourvo/users/we# (here a `home collection' could be just a collection which is given to us by our VO). We can get a list of entries in this collection from any Bartender.

\begin{enumerate}
\item We need to find a Bartender. Maybe we have a cached list of recently used Bartenders or we can get one from the information system. When we have an endpoint reference of a Bartender, we could call its list method with the LN \verb#/ourvo/users/we#.
\item The Bartender has to find a Librarian service, again using its cache of recently used Librarian services or get a new one from the information system. When the Bartender has an endpoint reference of a Librarian service, it could ask the Librarian to traverse the LN \verb#/ourvo/users/we#.
\item The Librarian needs the A-Hash service to access the stored data, when it has the endpoint reference of the A-Hash service, it could get the information about the root collection, which contains the GUID of the \verb#ourvo# sub-collection. Then the Librarian gets the entries of this \verb#ourvo# collection, and in it it can find the GUID of \verb#users#, and in the entries of \verb#users# there is the GUID of \verb#we#, which the Librarian returns to the Bartender with all the metadata.
\item The Bartender now has the GUID and the metadata of the collection \verb#/ourvo/users/we#, including the list of its entries. This is returned to us.
\item So we get the list of our \verb#/ourvo/users/we# collection, and now we realize that the file we want has the LN \verb#/ourvo/users/we/thefilewewant# and we know the GUID of it as well: e.g. \verb#a4b2e#. (Of course we know the GUID of the \verb#/ourvo/users/we# collection too, which is e.g. \verb#13245# and using this we could refer to our file as \verb#13245/thefilewewant# which means the entry called \verb#thefilewewant# in the collection with a GUID \verb#13245#.) We connect a Bartender again (the same one or maybe another one) to get the file with any of these LNs, the \verb#a4b2e# is the fastest solution because the Bartender need not to look up the whole LN again in the Librarian, a well-written client API should use this. With the get request we give the Bartender the list of transfer protocols we are able to use.
\item The Bartender contacts the Librarian to get the locations of the replicas of this file.
\item The Librarian returns all the metadata of the requested LN.
\item The Bartender chooses a `location' which consists of the ID of a Shepherd, and the referenceID of the file within the Shepherd. Using the information system or its local cache it could get the endpoint reference of the Shepherd. The Bartender initiates a transfer by the Shepherd, if the Shepherd supports one of the transfer protocols we give, it can create a transfer URL (TURL) with a protocol we can download.
\item The Shepherd returns the TURL to the Bartender.
\item The Bartender returns the TURL to us.
\item Now we have a TURL from which we can download it.
\end{enumerate}

@@

We have a file on our local disk we want to upload to a collection called \verb#/ourvo/common/docs#. (See Figure~\ref{fig:uploading}.)
\begin{enumerate}
\item We contact a Bartender to put the file, we give the size and checksum and other metadata, and the transfer protocols we want to use. And of course we give the Logical Name we want to be the name of the file, which in this case will be \verb#/ourvo/common/docs/proposal.pdf#
\item The Bartender ask a Librarian to traverse this LN.
\item If the Librarian can traverse the whole LN then this LN is already exists, but if the LN is still available and the parent exists, the Librarian’s response contains the GUID and metadata of the parent collection.
\item Then the Bartender creates a new file entry within the Librarian with all the information we gave. 
\item The Librarian returns the GUID of this new entry.
\item Then the Bartender add the name \verb#proposal.pdf# and the new GUID to the collection \verb#/ourvo/common/docs# and from now on there will be a valid LN \verb#/ourvo/common/docs/proposal.pdf# which points to a file which has no replica at all. If someone tried to download the file called \verb#/ourvo/common/docs/proposal.pdf# now, would get an error message `try again later'.
\item The Bartender (using the information system) chooses a Shepherd and gets its endpoint reference. Then the Bartender initiates uploading of the file to the Shepherd: the request includes the size and checksum of the file, the GUID, and the protocols we are able to use.
\item The Shepherd creates a transfer URL and a referenceID for this file and registers the GUID of the file in its own database and reports to the Librarian that there is a new replica with state \verb#creating#. The Librarian gets the message from the Shepherd and creates a new entry in the locations list of the given file with the serviceID and the referenceID the Shepherd have just reported. If someone tries to download this file now, still gets a `try again later' error message, because this new replica is still not \verb#alive#.
\item The Shepherd returns the the TURL to the Bartender.
\item The Bartender returns the TURL to us.
\item Then we can upload the file to this TURL.
\item The Shepherd detects that the file is arrived and reports the change of state to \verb#alive# to the  Librarian who alters the state in the given file-entry. At this point the file has only one replica. 
\end{enumerate}
\begin{itemize}
\item The Shepherd periodically checks the Librarian if this is less than the needed replica number, and if it is then it initiates creating a new replica by a Bartender.
\item The Bartender chooses another Shepherd, initiates the transfer then returns the TURL to the  first Shepherd which uploads the file to the new Shepherd.
\item All the Shepherds check periodically whether their files have enough replica, and if any of them find that there is more replica needed, it initiates creating a new. If more than one Shepherd of course could cause that there will be more replicas than needed. If a Shepherd finds out that a file has more replicas than needed it notifies a Bartender about it.
\item The Bartender ask the Librarian about all Shepherds this file has replicas on, flags this file as `removing a replica' which prevents other Bartenders to remove an other replica accidentally, then make a decision of which one is to be removed, then contacts the chosen Shepherd and asks it to remove the replica. The Shepherd then notify the Librarian, and the Librarian removes the replica, and removes the flag `removing a replica' as well.
\item If we cannot upload the file to the given TURL for some reason, we should remove the file entry from the collection, or we should call \verb#addReplica# to get a new TURL without removing and recreating the file.
\end{itemize}

@@

\begin{figure}[ht]
\centering{{\scalebox{0.7}{\includegraphics{arc1-storage-removing.pdf}}}
\caption{\label{fig:removing}Removing a file} }
\end{figure}
\begin{enumerate}
    \item If we want to remove a file, we should connect to a Bartender with the LN of the file we want to remove.
    \item The Bartender asks the Librarian to traverse the LN and return the list of parent collections of this file.
    \item The Librarian returns the data.
    \item If the file has only parent, it asks the Librarian to remove the file entry itself, and the entry from the parent collection. If the file has more parent collections (hard links) the Bartender only removes the entry from the parent collection and removes the parent from the list of parent  collections of this file.
    \item Next time the Shepherd does its periodic check, it asks the Librarian about each of its stored replicas, and finds out that one of them no longer exists, so it removes the replica from the storage node.
\end{enumerate}

@@

The services are written in Python and running in the HED\footnote{The ARC container - https://www.knowarc.eu/documents/Knowarc\_D1.2-2\_07.pdf} hosting environment. The HED itself is written in C++, but there are language bindings which allow us to write services in other languages, e.g. in Python or Java. The source tree of the storage services are in the NorduGrid Subversion repository\footnote{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/storage}.

The current version of the prototype has no information system and no security, these are soon to be integrated to the system.

The information system is needed to discover services, and to translate \emph{serviceID}s to endpoint references (URLs). Currently the URLs are written in the configuration files, and the Shepherd services are reporting their URLs to the Librarian, so the Bartender could ask for all alive Shepherds.

The security is needed to do proper authorization of the users, and to manage access policies of files and collections. ARC has its own policy language, for each file and collection there will be a policy XML document stored as a metadata. The storage services will use the properties extracted from the communication channel and these policies to make authorization decisions. If the properties and the policies are present, the decision is actually made by the security framework of HED.
